---
title: "Assignment 6"
author: "Kort Alexander and Alia Ajina"
date: "2023-05-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(sensitivity)
library(deSolve)
```

```{r}
# source function
source(here("R", "dgrowth.R"))

# set initial parameters and starting C
parms <- list(r = 0.01, K = 250, closer = 50, g = 2)
initial_C <- 10
time <- 1:300

# run ODE
results <- ode(y = initial_C, times = time, func = dgrowth, parms = parms)
colnames(results) = c("time", "C")
results_df <- as.data.frame(results)

# plot results
ggplot(results_df, aes(x = time, y = C)) +
  geom_line(size = 0.9, color = "darkgreen") +
  labs(x = "Years", y = "Forest Size (kg C)") + 
  theme_gray() + 
  scale_x_continuous(limits=c(0,300), expand = c(0,0), breaks = seq(0,300, by = 50)) + 
  scale_y_continuous(limits=c(0,200), expand = c(0,0), breaks = seq(0,200, by = 50))
```

# Sensitivity analysis

```{r}
# define parameter distributions and create samples for sobol object
sd_prop = 0.1
n_samp = 100

r = rnorm(mean = 0.01, sd = sd_prop * 0.01, n = n_samp)
closer = rnorm(mean = 50, sd = sd_prop * 50, n = n_samp)
K = rnorm(mean = 250, sd = sd_prop * 250, n = n_samp)
g = rnorm(mean = 2, sd = sd_prop * 2, n = n_samp)
X1 = cbind.data.frame(r=r, closer=closer, K=K, g=g)

r = rnorm(mean = 0.01, sd = sd_prop * 0.01, n = n_samp)
closer = rnorm(mean = 50, sd = sd_prop * 50, n = n_samp)
K = rnorm(mean = 250, sd = sd_prop * 250, n = n_samp)
g = rnorm(mean = 2, sd = sd_prop * 2, n = n_samp)
X2 = cbind.data.frame(r=r, closer=closer, K=K, g=g)

# create sobol object
sens_forest = sobolSalt(model = NULL,X1, X2, nboot = 300)
colnames(sens_forest$X) = c("r", "closer", "K", "g")

# define wrapper function to call ode and return carbon time series
wrapper = function(r, closer, K, g, Cinitial, times, func) {
  parms = list(r=r, closer=closer, K=K, g=g)
  result = ode(y=Cinitial, times=times, func=func, parms=parms)
  result=as.data.frame(result)
  colnames(result) = c("time", "C")
  max_C = max(result$C)
  return(list(max_C = max_C))
}

# set initial C and time and run wrapper function over our sobol object
times = 1:300
Cinitial = 10
allresults = as.data.frame(sens_forest$X) %>% pmap(wrapper, Cinitial = Cinitial, times = times, func = dgrowth)

# extract results into data frame
allres = allresults %>% map_dfr(`[`,c("max_C"))

# plot boxplot
ggplot(allres, aes(y = max_C)) +
  geom_boxplot(size = 0.9, color = "darkgreen") +
  labs(y = "Maximum carbon after 300 years") +
  theme_gray() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())

# get sobol indices
sens_forest_maxC = tell(sens_forest, allres$max_C)
rownames(sens_forest_maxC$T) = c("r", "closer", "K", "g")
sens_forest_maxC$T
rownames(sens_forest_maxC$S) = c("r", "closer", "K", "g")
sens_forest_maxC$S

```

Our results show that variability in g, K, and r have the greatest effect on variability in maximum forest size after 300 years. Climate change may influence g and r, the exponential and linear forest growth rates. Greater temperature and atmospheric carbon concentration will influence these growth rates depending on the forest species and location/current climate. For instance, some areas may see an initial increase in their g and r, until they reach a threshold where temperature or atmospheric carbon concentration impede on the forest growth rates.
