---
title: "Assignment 7"
author: "Kort Alexander and Alia Ajina"
date: "2023-05-30"
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: journal
    code_folding: show
    smooth_scroll: yes
    collapsed: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(deSolve)
library(ggpubr)
```

## Testing function without hunting

```{r}
source(here("R", "lotv_w_hunting.R"))

# Test function without hunting

# Initial conditions
currpop = c(prey=100, pred=10)

# Time points to see results
days = seq(from=1, to=200, by=1)

# Set parameters
pars = c(rprey=0.95, alpha=0.01, eff=0.6,pmort=0.4, K=2000, hunted_prop=0, min_pop = 1000)

# Run the model
res = ode(func=lotv_w_hunting, y=currpop, times=days, parms=pars)

# Visualize results
resl = as.data.frame(res) %>% pivot_longer(-time, names_to="animal", values_to="pop")
p1=ggplot(resl, aes(time, pop, col=animal))+ geom_line() + 
  labs(x = "Time (days)", y = "Species Population")
p1
```

## Testing with set parameters

```{r}

# Set parameters
pars = c(rprey=0.95, alpha=0.01, eff=0.6,pmort=0.4, K=2000, hunted_prop=0.1, min_pop = 100)

# Run the model
res = ode(func=lotv_w_hunting, y=currpop, times=days, parms=pars)

# Visualize results
resl = as.data.frame(res) %>% pivot_longer(-time, names_to="animal", values_to="pop")
p1=ggplot(resl, aes(time, pop, col=animal))+ geom_line() + labs(x = "Time (days)", y = "Species Population")
p1
```

## Comparing our stability metric

```{r}
# Creating our wrapper function
lotv_wrapper = function(currpop, days, rprey, alpha, eff, pmort, K, hunted_prop, min_pop) {
  parms = list(rprey=rprey, alpha=alpha, eff=eff, pmort=pmort, K=K, hunted_prop=hunted_prop, min_pop=min_pop)
  result = ode(y=currpop, times=days, func=lotv_w_hunting, parms=parms)
  colnames(result)=c("time","prey","pred")
  result_df = as.data.frame(result)
  if(result_df[length(result_df$time), 2] > 65)
    stable = TRUE
  else
    stable = FALSE
  return(stable)
}

# run function with hunting
hunted_prop <- seq(from = 0, to = 0.1, by = 0.025)
pars = list(rprey=0.95, alpha=0.01, eff=0.6,pmort=0.4, K=2000, hunted_prop = hunted_prop, min_pop = 100)
all_res = as.data.frame(pars) %>% pmap(lotv_wrapper, currpop = currpop, days = days)

# printing our data
data.frame(hunting = hunted_prop, stable = unlist(all_res))
```
## Defining our stability

Definition of stability: We defined a prey population as stable if it remains above 65 after 200 days. This may appear to be an arbitrary number, but this is the value where hunting leads to an unstable population, and no hunting leads to a stable population.

## Conclusions

For this species population, hunting permits will not be aloud. Due to the small population numbers, and it's vulnerability to hunting, if hunting permits were allotted, the population would dip below it's stability metric, and no longer be stable.
